An exact Riemann solver for one-dimensional multi-material elastic-plastic flows with Mie-Gr\"uneisen equation of state

### 2019-06-28 Manuscript revise 20190612.tex
222

For the elastic-plastic flow with the hypo-elastic constitutive model and  the von Mises' yielding condition,  some approximate Riemann solvers\cite{gavrilyuk2008modelling,cheng2015high,cheng2016harten,liumulti}  {\color{red}have been developed recently. However, for the exact Riemann solver, the research is relatively few, and them are only considered with  simple constitutive models and simple EOSs.} For example, Garaizar \cite{garaizar1991solution} and Miller \cite{miller2004iterative} introduced  an exact Riemann solver for elastic or hyper-elastic materials, Gao and Liu \cite{gao20171d,gao2018complete} firstly considered the yielding effect and developed an exact elastic-perfectly plastic solid Riemann solver. But there are three shotcomings in their works. {\color{red}Firstly, the EOS used by them is very simple: in the elastic state, the material derivative of pressure is the linear function of the strain rate, which means the energy equation is not necessary; in the plastic state,  a linear  "stiffened-gas" EOS was used.  Secondly, the states of elastic and plastic is using with very different equations systems and parameters, the seperation  deduces to more non-physical wave-structures cases in which the plastic wave  may be faster than the elastic wave. Thirdly, only the changing from  elastic to plastic is considered in their works, however, in fact  through an unloading and reloading process the plastic material also can change to elastic.}

@@

In this paper, we want to construct an exact Riemann solver for the system of 1D elastic-plastic flows with the Mie-Gr\"uneisen EOS, the hypo-elastic constitutive model and the von Mises' yielding condition. As every knows, the Mie-Gr\"uneisen EOS is complex and is used widely  in many engineering problems, which makes the energy equation must be included in the system of elastic-plastic flows and so we have to use some special methods to resolve the rarefaction wave. Otherwise, we analyse the Jacobian matrix and find, in the wave structures of Riemann solvers, there may be three to five waves, including one contact wave and other two to four non-linear waves. These nonlinear waves may be elastic shock waves, elastic rarefaction waves,  plastic shock waves or plastic rarefaction waves. Moreover, elastic wave runs always faster than the followed plastic wave. So there are thirty-six possible cases of the wave structures in the Riemann solver. {\color{red} Both the  changing from elastic to plastic and  the converse changing from plastic to elastic are considered.}
%For a given case, we will obtain theoretical relations of all the states as functions of density. At last,  through an iteration process of densities at both sides of the contact wave, the  exact solution will be obtained. Also, the half Riemann problem and its exact solution are considered in this paper.

@@

%%The governing equations system is given as
%% \begin{equation}\label{eq:1d}
%   \left\{ \begin{aligned}
%       & \partial _t \rho +\partial_x(\rho u)=0,\\
%       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
%       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
%       &\partial _t s_{xx}+u\partial_xs_{xx}-\frac{4}{3}\partial_x u=0,\\
%& |s_{xx}|\leq \frac{2}{3}Y_{0}, \\
%        \end{aligned}
%  \right.
%\end{equation}
%It contains the following parts.
\subsection{Motion equations}
For  a continuous one-dimensional homogeneous solid, the motion equations  in differential form are
\begin{equation*}
\partial_t \mathbf{{U}} + \partial _x \bm{F}(\mathbf{{U}}) = 0, \   x \in \   \Omega \subset \mathbf{R}, \  t>0,
\end{equation*}
where
\begin{equation}
  \mathbf{U} = \left[ \begin{array}{l}
      \rho \\
      \rho u \\
      \rho  E \\
    \end{array}
  \right],
  \hspace{0.3cm}
  \mathbf{F} = \left[ \begin{array}{l}
      \rho u \\
      \rho u^2 -\sigma\\
      (\rho E -\sigma)u\\
  \end{array} \right],
\end{equation}
$\rho$, $u$, $\sigma$ and $E$ are  the density, velocity in $x-$direction, Cauchy stress and total energy per unit volume, respectively, $E$ has the relation with specific internal energy $e$ as
\begin{equation} \label{te1}
  E = e+\frac{1}{2}u^2,
\end{equation}
\begin{equation}\label{sigma1}
  \sigma = -p +s_{xx},
\end{equation}
where $p$ and $s_{xx}$ denote hydrostatic pressure and deviatoric stress in the $x-$ direction, respectively.

@@

\subsection{The yielding condition}
The Von Mises' yielding condition is used here to describe the elastic limit. In one spatial dimension, the von Mises' yielding criterion is given by
\begin{equation}
  |s_{xx}| \le \frac{2}{3}Y_0,
\end{equation}
where $Y_0$ is the yield strength of the material in simple tension.

@@

The Riemann problem for the 1D time dependent elastic-plastic equations is given as follows:
 \begin{equation}\label{eq:1d}
   \left\{ \begin{aligned}
       & \partial _t \rho +\partial_x(\rho u)=0,\\
       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
       &\partial _t s_{xx}+u\partial_xs_{xx}-\frac{4}{3}\partial_x u=0,\\
& |s_{xx}|\leq \frac{2}{3}Y_{0}, \\
       &Q(x,t = 0) = \left\{\begin{aligned}
           Q_L, \hspace{0.1cm} \text{if} \hspace{0.1cm} x<0, \\
           Q_R, \hspace{0.1cm} \text{if} \hspace{0.1cm} x\ge 0, \\
       \end{aligned}\right.
     \end{aligned}
  \right.
\end{equation}
where $Q = (\rho, \rho u, \rho E, s_{xx})^T$.

The above equations are just for the elastic status of the solid material.
If the material are in the plastic state, the above equations can be simplified
and correspondingly the jacobian matrix and sonic velocity are different.
We will discuss them separately.
\subsection{Elastic state}
\subsubsection{Jacobian matrix in elastic regions} %and the eigenvalues and eigenvectors}
For the Mie-Gr\"uneisen EOS, if the material is not yielding,
\begin{equation}
  |s_{xx}| < \frac{2}{3}Y_0,
\end{equation}
the system (\ref{eq:1d}) can be written as
\begin{equation}
 \partial_t \mathbf{Q} +\mathbf{J}_{e}(\mathbf{{Q}})\partial_x\mathbf{Q} = 0,
\end{equation}
where $Q = (\rho, \rho u, \rho E, s_{xx})$, and the Jacobian matrix is
\begin{equation}\label{eq:Jcb}
  \mathbf{J}_{e}(Q) = \left[\begin{array}{llll}
      0 & 1 & 0 & 0 \\
      -u^2 + \frac{\partial p}{\partial \rho} +\Gamma(\frac{u^2}{2}-e)& u(2-\Gamma)& \Gamma & -1 \\
	  (\Gamma(\frac{u^2}{2}-e)-e-\frac{u^2}{2}+\frac{\sigma}{\rho}+\frac{\partial p}{\partial \rho})u & -\Gamma u^2 -\frac{\sigma}{\rho}+\frac{u^2}{2} +e & (1+\Gamma)u& -u\\
    \frac{4}{3}\mu\frac{u}{\rho} & -\frac{4}{3}\mu\frac{1}{\rho}& 0 & u \\
\end{array}
\right],
\end{equation}
where $\Gamma = \frac{\Gamma_0\rho_0}{\rho} $.

The eigenvalues of $\mathbf{J}_{e}(\mathbf{Q})$ are given as
\begin{equation}
  \lambda_1 =\lambda_2 = u, \hspace{0.3cm} \lambda_3 = u-c_{e}, \hspace{0.3cm} \lambda_4 = u+c_{e},
\end{equation}
where $c_e$ means the sonic speed of the solid in the elastic state,
\begin{equation}\label{eq:c_e}
  \left\{ \begin{aligned}
      & c_e = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx} +\frac{4}{3}\frac{\mu}{\rho}},\\
    &   a^2 = \frac{\partial p}{\partial \rho} + \frac{p}{\rho^2}\frac{\partial p}{\partial e} = a^2_0 \frac{\partial f}{\partial \eta} + \frac{p}{\rho^2}\rho_0 \Gamma_0.
      \end{aligned} \right.
    \end{equation}
The corresponding right eigenvectors are
\begin{equation}\label{eq:eiv}
  r_1 = \left[ \begin{array}{l}
      \frac{1}{b_1} \\
      \frac{u}{b_1} \\
      0 \\
      1 \\
    \end{array}
    \right], \hspace{0.2cm}
    r_2= \left[ \begin{array}{l}
        -\frac{\Gamma}{b_1} \\
        -\frac{\Gamma u}{b_1} \\
        1 \\
        0\\
      \end{array}
    \right], \hspace{0.2cm}
r_3 =   \frac{1}{\phi^2}\left[\begin{array}{l}
        1 \\
        u-c_e \\
        h -uc_e \\
        \phi^2
      \end{array}
    \right], \hspace{0.2cm}
r_4 = \frac{1}{\phi^2}\left[\begin{array}{l}
        1 \\
        u+c_e \\
        h +uc_e \\
        \phi^2
      \end{array}
    \right],
  \end{equation}
  where
  \begin{equation}
    b_1 = \frac{\partial p}{\partial \rho} - \Gamma E,  \quad h = E +\frac{p-s_{xx}}{\rho},
  \end{equation}
  and
  \begin{equation}
    \phi^2 = a^2 -\frac{\rho_0}{\rho^2} \Gamma_0 s_{xx}-c_e^2 = -\frac{4\mu}{3}\frac{1}{\rho}.
  \end{equation}

@@

Thanks to (\ref{eq:mt}), the equations of the density and the deviatoric stress in Eq.(\ref{eq:1d}) can be written as
  \begin{equation}\label{eq:d1}
    \frac{\partial u}{\partial x} = -\frac{1}{\rho}\frac{d\rho}{dt},
  \end{equation}
  and
  \begin{equation}\label{eq:s1}
    \frac{ds_{xx}}{dt}=\frac{4}{3}\mu\frac{\partial u}{\partial x}.
  \end{equation}

  Substituting (\ref{eq:d1}) into (\ref{eq:s1}) yields
%  We can get a relation of density and deviatoric stress,
  \begin{equation}
    \frac{ds_{xx}}{dt}=-\frac{4}{3}\mu \frac{1}{\rho}\frac{d\rho}{dt}.
\end{equation}
Integrate the above equation from the data in front of a wave to the data behind the wave and perform some simple algebraic manipulations, one can get
\begin{equation}\label{eq:rhosxx}
  s_{xx}+\frac{4}{3}\mu\text{ln}(\rho) = \text{constant}
\end{equation}
This relation always hold in the elastic state.

\subsubsection{Relations across the contact wave}\label{sec:contacte}
  For  a  system without molecular diffusion, there is no materials convecting  across the contact wave or interface, so the velocities on two sides of  the discontinuity are always equal. % .
  This can also be verified by the eigenvectors  in Eq.(\ref{eq:eiv}) and  Eq.(\ref{eq:eivp}).

@@

Thanks to Eq.(\ref{eq:eiv}), for the $\lambda_{1}$-wave we
have
\begin{equation}   \label{e23a}
\frac{d \rho}{\frac{1}{b_{1}}} = \frac{d \rho u}{\frac{u
}{b_{1}}}=\frac{d \rho E}{0} = \frac{d s_{xx}}{1}.
\end{equation}
From the above equations, we can easily deduce that
\begin{equation}   \label{e23b}
du = 0, \quad d(s_{xx}-p)=0,
\end{equation}
 which means
\begin{equation}   \label{e23c}
  u_{L}^{\ast}=u_{R}^{\ast},
\end{equation}
and
\begin{equation}   \label{e23d}
\sigma_{x,L}^{\ast}=\sigma_{x,R}^{\ast},
\end{equation}
where $()_{L}^{\ast}$ and $()_{R}^{\ast}$ denote $()$ in the region
of $\mathbf{Q}_{L}^{\ast}$ and $\mathbf{Q}_{R}^{\ast}$,
respectively. Here we do not show the details of the derivation for a simple
 presentation.

@@

By using the same methods for (\ref{eq:RH3}), (\ref{eq:RH3}) can be written as
\begin{equation}
  t(u_1-u_2)(E_2-E_1) =\sigma_2 u_2-\sigma_1u_1.
\end{equation}
Because of $E= e+\frac{1}{2}u^2$, we can get
\begin{equation}\label{eq:e2e1}
  e_2 - e_1 = - \frac{\sigma_1 +\sigma_2}{2t}.
\end{equation}
Using the EOS of Mie-Gr\"uneisen (\ref{eq:mie}), can get
\begin{equation}
  e = c_0p -c_1f(\rho/\rho_0),
\end{equation}
where $c_0 = \frac{1}{\rho_0\Gamma_0}$ and $c_1 = \frac{a_0^2}{\Gamma_0}$.
Put the above equation into (\ref{eq:e2e1}), we can solve the pressure $p_2$ out as a function of $\rho_2$.
\begin{equation}\label{eq:shocke}
  p_2= \frac{2t(c_1f(\rho_2/\rho_0)+e_1)-(\sigma_1+s_{xx2})}{2tc_0-1}.
\end{equation}

@@

When the material is yielding,
\begin{equation}\label{eq:sxxp}
  |s_{xx}| = \frac{2}{3}Y_0,
\end{equation}
the equations of Riemann problem  can be written as
\begin{equation}\label{eq:1dp}
   \left\{ \begin{aligned}
       & \partial _t \rho +\partial_x(\rho u)=0,\\
       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
& |s_{xx}| = \frac{2}{3}Y_{0}, \\
       &U(x,t = 0) = \left\{\begin{aligned}
           U_L, \hspace{0.1cm} \text{if} \hspace{0.1cm} x<0, \\
           U_R, \hspace{0.1cm} \text{if} \hspace{0.1cm} x\ge 0, \\
       \end{aligned}\right.
     \end{aligned}
  \right.
\end{equation}
where $\mathbf{U} = (\rho, \rho u, \rho E )$.
  \subsubsection{Jacobian matrix in  plastic  regions}
  %\footnote{\href{https://nbviewer.jupyter.org/github/bfly123/Exact-Elastic-plastic-Riemann-Solver-/blob/master/\%20CharacteristicForPlasticStateAfterYielding.ipynb\#}{Code site link}}}
 The motion equations of (\ref{eq:1dp}) can be written as
\begin{equation}
  \partial_t \mathbf{{U}} + \mathbf{J}_p(\mathbf{U})\partial_x \mathbf{{U}}= 0,
\end{equation}
 where the Jacobian matrix is
\begin{equation}
\mathbf{J}_p(\mathbf{U}) = \left[\begin{array}{lll}
      0 & 1 & 0   \\
      -u^2 + \frac{\partial p}{\partial \rho} +\Gamma(\frac{u^2}{2}-e)& u(2-\Gamma)& \Gamma \\
	  (\Gamma(\frac{u^2}{2}-e)-e-\frac{u^2}{2}+\frac{\sigma}{\rho}+\frac{\partial p}{\partial \rho})u +\frac{u^2}{2} & -\Gamma u^2 -\frac{\sigma}{\rho} +e & (1+\Gamma)u\\
\end{array}
\right].
\end{equation}

The eigenvalues of $\mathbf{J}_p(\mathbf{Q})$ are given as
$$\lambda_1 = u,\quad \lambda_2 = u-c_p, \quad \lambda_3 = u+c_p,$$
where
\begin{equation}\label{eq:c_p}
   c_p = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}}.
\end{equation}
%\begin{equation}\label{eq:c_p}
%  \left\{ \begin{aligned}
%	  & c_p = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}} ,\\
%    &   a^2 = \frac{\partial p}{\partial \rho} + \frac{p}{\rho^2}\frac{\partial p}{\partial e} = a^2_0 \frac{\partial f}{\partial \eta} + \frac{p}{\rho^2}\rho_0 \Gamma_0.
%      \end{aligned} \right.
%    \end{equation}
The corresponding right eigenvectors are
\begin{equation}\label{eq:eivp}
  r_1 = \left[ \begin{array}{l}
	  -\frac{\Gamma}{b_1} \\
	  -\frac{\Gamma u}{b_1} \\
	  1\\
  \end{array}\right], \quad
  r_2 = \frac{1}{h-uc_p} \left[  \begin{array}{l}
	  1 \\
	  u-c_p \\
	  h-uc_p\\
  \end{array}\right], \quad
 r_3 = \frac{1}{h+uc_p}\left[ \begin{array}{l}
	  1 \\
	  u+c_p \\
	  h+uc_p\\
  \end{array}\right].
\end{equation}
%where
% \begin{equation}
%    b_1 = \frac{\partial p}{\partial \rho} - \Gamma E,  \quad h = E +\frac{p-s_{xx}}{\rho}.
%  \end{equation}

@@

Now we consider the constructing details of the exact  Riemann  solver. For  the Riemann problem in Section \ref{sec:riemann}, there are  $6\times 6$ possible  cases in the Riemann solution with different wave structures.  The left six cases are  shown in Fig.\ref{fig:cases}.

@@

	Given $\rho _{L(k)}^*$ and a  $\rho _{R(k)}^*$ in the $k$-th  iteration step, we can use the methods introduced in Section \ref{sec:deter} to determine the case of wave structure of this Riemann problem. In this procedure for solving this Riemann problem, the subscript $(k)$ means the variable in the $k$-th  iteration step.

@@

  After determining the structure case, we need  to solve the Cauchy stresses and the velocities in regions $\mathbf{Q}^*_L$ and $\mathbf{Q}^*_R$, those are given in Section \ref{sec:functions}.

>>>>

For the elastic-plastic flow with the hypo-elastic constitutive model and  the von Mises' yielding condition,  some approximate Riemann solvers\cite{gavrilyuk2008modelling,cheng2015high,cheng2016harten,liumulti}  {\color{red}have been developed recently. However, for the exact Riemann solver, the research is relatively few, and them are only considered with  simple constitutive models and simple EOSs.} For example, Garaizar \cite{garaizar1991solution} and Miller \cite{miller2004iterative} introduced  an exact Riemann solver for elastic or hyper-elastic materials, Gao and Liu \cite{gao20171d,gao2018complete} firstly considered the yielding effect and developed an exact elastic-perfectly plastic solid Riemann solver. But there are three issues need to improve or overcome in their works. {\color{red}Firstly, the EOS used by them is very simple: in the elastic state, the material derivative of pressure is the linear function of the strain rate, which means the energy equation is not necessary; in the plastic state,  a linear  "stiffened-gas" EOS was used.  Secondly, the states of elastic and plastic is using with very different equations systems and parameters, the seperation  deduces to more possible non-physical wave-structure cases in which the plastic wave  may run faster than the elastic wave. Thirdly, only the changing from  elastic to plastic is considered in their works, however, in fact  through an unloading and reloading process the plastic material  can also change to elastic states.}

@@

In this paper, we want to construct an exact Riemann solver for the system of 1D elastic-plastic flows with the Mie-Gr\"uneisen EOS, the hypo-elastic constitutive model and the von Mises' yielding condition. As every knows, the Mie-Gr\"uneisen EOS is complex and is used widely  in many engineering problems, which makes the energy equation must be included in the system of elastic-plastic flows and so we have to use some special methods to resolve the rarefaction wave. Otherwise, we analyse the Jacobian matrix and find, in the wave structures of Riemann solvers, there may be three to five waves, including one contact wave and other two to four non-linear waves. These nonlinear waves may be elastic shock waves, elastic rarefaction waves,  plastic shock waves or plastic rarefaction waves. Moreover, elastic wave always runs faster than the followed plastic wave. So there are thirty-six possible cases of the wave structures in the Riemann solver. {\color{red} Besides, both the changing from elastic to plastic and  the converse changing from plastic to elastic are considered in the Riemann problem.}
%For a given case, we will obtain theoretical relations of all the states as functions of density. At last,  through an iteration process of densities at both sides of the contact wave, the  exact solution will be obtained. Also, the half Riemann problem and its exact solution are considered in this paper.

@@

%%The governing equations system is given as
%% \begin{equation}\label{eq:1d}
%   \left\{ \begin{aligned}
%       & \partial _t \rho +\partial_x(\rho u)=0,\\
%       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
%       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
%       &\partial _t s_{xx}+u\partial_xs_{xx}-\frac{4}{3}\partial_x u=0,\\
%& |s_{xx}|\leq \frac{2}{3}Y_{0}, \\
%        \end{aligned}
%  \right.
%\end{equation}
%It contains the following parts.
\subsection{Motion equations}
For  a continuous one-dimensional homogeneous solid, the motion equations  in  the differential form are
\begin{equation*}
\partial_t \mathbf{{U}} + \partial _x \bm{F}(\mathbf{{U}}) = 0, \   x \in \   \Omega \subset \mathbf{R}, \  t>0,
\end{equation*}
where
\begin{equation}
  \mathbf{U} = \left[ \begin{array}{l}
      \rho \\
      \rho u \\
      \rho  E \\
    \end{array}
  \right],
  \hspace{0.3cm}
  \mathbf{F} = \left[ \begin{array}{l}
      \rho u \\
      \rho u^2 -\sigma\\
      (\rho E -\sigma)u\\
  \end{array} \right],
\end{equation}
$\rho$, $u$, $\sigma$ and $E$ are  the density, velocity in $x-$direction, Cauchy stress and total energy per unit volume, respectively, $E$ has the relation with the  specific internal energy $e$ as
\begin{equation} \label{te1}
  E = e+\frac{1}{2}u^2,
\end{equation}
\begin{equation}\label{sigma1}
  \sigma = -p +s_{xx},
\end{equation}
where $p$ and $s_{xx}$ denote the  hydrostatic pressure and the deviatoric stress in the $x-$ direction, respectively.

@@

\subsection{The yielding condition}
The Von Mises' yielding condition is used here to describe the elastic limit. In one spatial dimension, the von Mises' yielding criterion is given by
\begin{equation}
  |s_{xx}| \le \frac{2}{3}Y_0,
\end{equation}
where $Y_0$ is the yield strength of the material in  simple tensions.

@@

The Riemann problem for 1D time dependent elastic-plastic equations is given as follows:
 \begin{equation}\label{eq:1d}
   \left\{ \begin{aligned}
       & \partial _t \rho +\partial_x(\rho u)=0,\\
       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
       &\partial _t s_{xx}+u\partial_xs_{xx}-\frac{4}{3}\partial_x u=0,\\
& |s_{xx}|\leq \frac{2}{3}Y_{0}, \\
       &Q(x,t = 0) = \left\{\begin{aligned}
           Q_L, \hspace{0.1cm} \text{if} \hspace{0.1cm} x<0, \\
           Q_R, \hspace{0.1cm} \text{if} \hspace{0.1cm} x\ge 0, \\
       \end{aligned}\right.
     \end{aligned}
  \right.
\end{equation}
where $Q = (\rho, \rho u, \rho E, s_{xx})^T$.

The above equations are just for the elastic status of the solid material.
If the materials are in the plastic state, the above equations can be simplified
and the jacobian matrix and sonic velocity are different,  correspondingly.
We will discuss them separately.
\subsection{Elastic state}
\subsubsection{Jacobian matrix in elastic regions} %and the eigenvalues and eigenvectors}
For the Mie-Gr\"uneisen EOS, if the material is not yielding,
\begin{equation}
  |s_{xx}| < \frac{2}{3}Y_0,
\end{equation}
the system (\ref{eq:1d}) can be written as
\begin{equation}
 \partial_t \mathbf{Q} +\mathbf{J}_{e}(\mathbf{{Q}})\partial_x\mathbf{Q} = 0,
\end{equation}
where $Q = (\rho, \rho u, \rho E, s_{xx})$, and the Jacobian matrix is
\begin{equation}\label{eq:Jcb}
  \mathbf{J}_{e}(Q) = \left[\begin{array}{llll}
      0 & 1 & 0 & 0 \\
      -u^2 + \frac{\partial p}{\partial \rho} +\Gamma(\frac{u^2}{2}-e)& u(2-\Gamma)& \Gamma & -1 \\
	  (\Gamma(\frac{u^2}{2}-e)-e-\frac{u^2}{2}+\frac{\sigma}{\rho}+\frac{\partial p}{\partial \rho})u & -\Gamma u^2 -\frac{\sigma}{\rho}+\frac{u^2}{2} +e & (1+\Gamma)u& -u\\
    \frac{4}{3}\mu\frac{u}{\rho} & -\frac{4}{3}\mu\frac{1}{\rho}& 0 & u \\
\end{array}
\right],
\end{equation}
where $\Gamma = \frac{\Gamma_0\rho_0}{\rho} $.

The eigenvalues of $\mathbf{J}_{e}(\mathbf{Q})$ are given as
\begin{equation}
  \lambda_1 =\lambda_2 = u, \hspace{0.3cm} \lambda_3 = u-c_{e}, \hspace{0.3cm} \lambda_4 = u+c_{e},
\end{equation}
where $c_e$ means the sonic speed of the solid in the elastic state,
\begin{equation}\label{eq:c_e}
  \left\{ \begin{aligned}
      & c_e = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx} +\frac{4}{3}\frac{\mu}{\rho}},\\
    &   a^2 = \frac{\partial p}{\partial \rho} + \frac{p}{\rho^2}\frac{\partial p}{\partial e} = a^2_0 \frac{\partial f}{\partial \eta} + \frac{p}{\rho^2}\rho_0 \Gamma_0.
      \end{aligned} \right.
    \end{equation}
 Corresponding right eigenvectors are
\begin{equation}\label{eq:eiv}
  r_1 = \left[ \begin{array}{l}
      \frac{1}{b_1} \\
      \frac{u}{b_1} \\
      0 \\
      1 \\
    \end{array}
    \right], \hspace{0.2cm}
    r_2= \left[ \begin{array}{l}
        -\frac{\Gamma}{b_1} \\
        -\frac{\Gamma u}{b_1} \\
        1 \\
        0\\
      \end{array}
    \right], \hspace{0.2cm}
r_3 =   \frac{1}{\phi^2}\left[\begin{array}{l}
        1 \\
        u-c_e \\
        h -uc_e \\
        \phi^2
      \end{array}
    \right], \hspace{0.2cm}
r_4 = \frac{1}{\phi^2}\left[\begin{array}{l}
        1 \\
        u+c_e \\
        h +uc_e \\
        \phi^2
      \end{array}
    \right],
  \end{equation}
  where
  \begin{equation}
    b_1 = \frac{\partial p}{\partial \rho} - \Gamma E,  \quad h = E +\frac{p-s_{xx}}{\rho},
  \end{equation}
  and
  \begin{equation}
    \phi^2 = a^2 -\frac{\rho_0}{\rho^2} \Gamma_0 s_{xx}-c_e^2 = -\frac{4\mu}{3}\frac{1}{\rho}.
  \end{equation}

@@

Thanks to (\ref{eq:mt}), equations of the density and the deviatoric stress in Eq.(\ref{eq:1d}) can be written as
  \begin{equation}\label{eq:d1}
    \frac{\partial u}{\partial x} = -\frac{1}{\rho}\frac{d\rho}{dt},
  \end{equation}
  and
  \begin{equation}\label{eq:s1}
    \frac{ds_{xx}}{dt}=\frac{4}{3}\mu\frac{\partial u}{\partial x}.
  \end{equation}

  Substituting (\ref{eq:d1}) into (\ref{eq:s1}) yields
%  We can get a relation of density and deviatoric stress,
  \begin{equation}
    \frac{ds_{xx}}{dt}=-\frac{4}{3}\mu \frac{1}{\rho}\frac{d\rho}{dt}.
\end{equation}
Integrating the above equation from the data in front of a wave to the data behind the wave and performing some simple algebraic manipulations, one can get
\begin{equation}\label{eq:rhosxx}
  s_{xx}+\frac{4}{3}\mu\text{ln}(\rho) = \text{constant}
\end{equation}
This relation always hold in the elastic state.

\subsubsection{Relations across the contact wave}\label{sec:contacte}
  For  a  system without molecular diffusion, there is no materials convecting  across the contact wave or interface, so the velocities on two sides of  the discontinuity are always equal. % .
  This can also be verified by eigenvectors in Eq.(\ref{eq:eiv}) and  Eq.(\ref{eq:eivp}).

@@

Thanks to Eq.(\ref{eq:eiv}), for the $\lambda_{1}$-wave we
have
\begin{equation}   \label{e23a}
\frac{d \rho}{\frac{1}{b_{1}}} = \frac{d \rho u}{\frac{u
}{b_{1}}}=\frac{d \rho E}{0} = \frac{d s_{xx}}{1}.
\end{equation}
From above equations, we can easily deduce that
\begin{equation}   \label{e23b}
du = 0, \quad d(s_{xx}-p)=0,
\end{equation}
 which means
\begin{equation}   \label{e23c}
  u_{L}^{\ast}=u_{R}^{\ast},
\end{equation}
and
\begin{equation}   \label{e23d}
\sigma_{x,L}^{\ast}=\sigma_{x,R}^{\ast},
\end{equation}
where $()_{L}^{\ast}$ and $()_{R}^{\ast}$ denote $()$ in the region
of $\mathbf{Q}_{L}^{\ast}$ and $\mathbf{Q}_{R}^{\ast}$,
respectively. Here we do not show the details of the derivation for a simple
 presentation.

@@

By using the same methods for (\ref{eq:u2u1}), (\ref{eq:RH3}) can be written as
\begin{equation}
  t(u_1-u_2)(E_2-E_1) =\sigma_2 u_2-\sigma_1u_1.
\end{equation}
Because of $E= e+\frac{1}{2}u^2$, we can get
\begin{equation}\label{eq:e2e1}
  e_2 - e_1 = - \frac{\sigma_1 +\sigma_2}{2t}.
\end{equation}
Using the EOS of Mie-Gr\"uneisen (\ref{eq:mie}), can get
\begin{equation}
  e = c_0p -c_1f(\rho/\rho_0),
\end{equation}
where $c_0 = \frac{1}{\rho_0\Gamma_0}$ and $c_1 = \frac{a_0^2}{\Gamma_0}$.
Put the above equation into (\ref{eq:e2e1}), we can solve the pressure $p_2$ out as a function of $\rho_2$.
\begin{equation}\label{eq:shocke}
  p_2= \frac{2t(c_1f(\rho_2/\rho_0)+e_1)-(\sigma_1+s_{xx2})}{2tc_0-1}.
\end{equation}

@@

When the material is yielding,
\begin{equation}\label{eq:sxxp}
  |s_{xx}| = \frac{2}{3}Y_0,
\end{equation}
the equations of Riemann problem  can be written as
\begin{equation}\label{eq:1dp}
   \left\{ \begin{aligned}
       & \partial _t \rho +\partial_x(\rho u)=0,\\
       & \partial _t (\rho u)+\partial_x(\rho u^2 + p -s_{xx})=0,\\
       &\partial _t (\rho E)+\partial_x\left[(\rho E + p -s_{xx})u\right]=0,\\
& |s_{xx}| = \frac{2}{3}Y_{0}, \\
       &U(x,t = 0) = \left\{\begin{aligned}
           U_L, \hspace{0.1cm} \text{if} \hspace{0.1cm} x<0, \\
           U_R, \hspace{0.1cm} \text{if} \hspace{0.1cm} x\ge 0, \\
       \end{aligned}\right.
     \end{aligned}
  \right.
\end{equation}
where $\mathbf{U} = (\rho, \rho u, \rho E )$.
  \subsubsection{Jacobian matrix in  plastic  regions}
  %\footnote{\href{https://nbviewer.jupyter.org/github/bfly123/Exact-Elastic-plastic-Riemann-Solver-/blob/master/\%20CharacteristicForPlasticStateAfterYielding.ipynb\#}{Code site link}}}
 Motion equations of (\ref{eq:1dp}) can be written as
\begin{equation}
  \partial_t \mathbf{{U}} + \mathbf{J}_p(\mathbf{U})\partial_x \mathbf{{U}}= 0,
\end{equation}
 where the Jacobian matrix is
\begin{equation}
\mathbf{J}_p(\mathbf{U}) = \left[\begin{array}{lll}
      0 & 1 & 0   \\
      -u^2 + \frac{\partial p}{\partial \rho} +\Gamma(\frac{u^2}{2}-e)& u(2-\Gamma)& \Gamma \\
	  (\Gamma(\frac{u^2}{2}-e)-e-\frac{u^2}{2}+\frac{\sigma}{\rho}+\frac{\partial p}{\partial \rho})u +\frac{u^2}{2} & -\Gamma u^2 -\frac{\sigma}{\rho} +e & (1+\Gamma)u\\
\end{array}
\right].
\end{equation}

Eigenvalues of $\mathbf{J}_p(\mathbf{Q})$ are given as
$$\lambda_1 = u,\quad \lambda_2 = u-c_p, \quad \lambda_3 = u+c_p,$$
where
\begin{equation}\label{eq:c_p}
   c_p = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}}.
\end{equation}
%\begin{equation}\label{eq:c_p}
%  \left\{ \begin{aligned}
%	  & c_p = \sqrt{a^2-\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}} ,\\
%    &   a^2 = \frac{\partial p}{\partial \rho} + \frac{p}{\rho^2}\frac{\partial p}{\partial e} = a^2_0 \frac{\partial f}{\partial \eta} + \frac{p}{\rho^2}\rho_0 \Gamma_0.
%      \end{aligned} \right.
%    \end{equation}
The corresponding right eigenvectors are
\begin{equation}\label{eq:eivp}
  r_1 = \left[ \begin{array}{l}
	  -\frac{\Gamma}{b_1} \\
	  -\frac{\Gamma u}{b_1} \\
	  1\\
  \end{array}\right], \quad
  r_2 = \frac{1}{h-uc_p} \left[  \begin{array}{l}
	  1 \\
	  u-c_p \\
	  h-uc_p\\
  \end{array}\right], \quad
 r_3 = \frac{1}{h+uc_p}\left[ \begin{array}{l}
	  1 \\
	  u+c_p \\
	  h+uc_p\\
  \end{array}\right].
\end{equation}
%where
% \begin{equation}
%    b_1 = \frac{\partial p}{\partial \rho} - \Gamma E,  \quad h = E +\frac{p-s_{xx}}{\rho}.
%  \end{equation}

@@

Now we consider the constructing details of the exact  Riemann  solver. For  the Riemann problem in Section \ref{sec:Riemann}, there are  $6\times 6$ possible  cases in the Riemann solution with different wave structures.  The left six cases are  shown in Fig.\ref{fig:cases}.

@@

	Given $\rho _{L(k)}^*$ and  $\rho _{R(k)}^*$ in the $k$-th  iteration step, we can use the method introduced in Section \ref{sec:deter} to determine the case of wave structure of this Riemann problem. In the  procedure for solving the  Riemann problem, the subscript $(k)$ means the variable in the $k$-th  iteration step.

@@

  After determining the structure case, we need  to solve  Cauchy stresses and velocities in regions $\mathbf{Q}^*_L$ and $\mathbf{Q}^*_R$, those are given in Section \ref{sec:functions}.
